Precise predictions for same-sign W-bosons scattering at the LHC

### 2018-02-07 Draft/distributionsNLO.tex
Minor changes on draft, mostly typos

Concerning the transverse momentum (top) and rapidity (bottom) of the hardest jet shown in Fig.~\ref{fig:distNLO2}, the situation is rather different.
While {\sc MadGraph5\_aMC@NLO} is very close to the full prediction for low transverse momentum, it is diverging from it at larger transverse momentum.
This is in contrast with ({\sc BONSAY} and {\sc POWHEG}) which approximate the full computation reasonably well over the whole range and in particular the the high transverse-momentum region.
Finally, {\sc VBFNLO} predicts higher rates over the whole range apart around $200\GeV$ where it is in perfect agreement with the complete calculation.
Concerning the rapidity of the hardest jet, {\sc VBFNLO} is in good agreement with {\sc MoCaNLO}+{\sc Recola} in the rapidity range $\pm3$.
For larger rapidity, the other codes constitute a better description of the full process at order $\mathcal{O}(\alpha_{\rm s}\alpha^6)$.

@@

The last set of differential distributions are the invariant mass of the two charged lepton (top) and the Zeppenfeld variable for the anti-muon (bottom).
Concerning the comparison of the predictions, both distributions display a rather similar behaviour.
Indeed, the hierarchy mentioned previously is here respected and enhanced towards high invariant mass or high Zeppenfeld variable.
{\sc MoCaNLO}+{\sc Recola} and {\sc VBFNLO} are in rather good agreement for both distributions for the kinematic range displayed here.
The other three VBS approximations seem to be close to each other.

>>>>

Concerning the transverse momentum (top) and rapidity (bottom) of the hardest jet shown in Fig.~\ref{fig:distNLO2}, the situation is rather different.
While {\sc MadGraph5\_aMC@NLO} is very close to the full prediction for low transverse momentum, it is diverging from it at larger transverse momentum.
This is in contrast with {\sc BONSAY} and {\sc POWHEG} which approximate the full computation reasonably well over the whole range and in particular the high transverse-momentum region.
Finally, {\sc VBFNLO} predicts higher rates over the whole range apart around $200\GeV$ where it is in perfect agreement with the complete calculation.
Concerning the rapidity of the hardest jet, {\sc VBFNLO} is in good agreement with {\sc MoCaNLO}+{\sc Recola} in the rapidity range $|y_{j_1}| < 3$.
For larger rapidity, the other codes constitute a better description of the full process at order $\mathcal{O}(\alpha_{\rm s}\alpha^6)$.

@@

The last set of differential distributions is the invariant mass of the two charged lepton (top) and the Zeppenfeld variable for the anti-muon (bottom).
Concerning the comparison of the predictions, both distributions display a rather similar behaviour.
Indeed, the hierarchy mentioned previously is here respected and enhanced towards high invariant mass or high Zeppenfeld variable.
{\sc MoCaNLO}+{\sc Recola} and {\sc VBFNLO} are in rather good agreement for both distributions for the kinematic range displayed here.
The other three VBS approximations seem to be close to each other.

### 2018-02-07 Draft/introduction.tex
small typos fixed
Minor changes on draft, mostly typos

In the last few years, several leading order (LO) and next-to-leading order (NLO) computations became available for both the VBS process~\cite{Jager:2009xx,Jager:2011ms,Denner:2012dz,Rauch:2016pai} and its QCD-induced irreducible background process~\cite{Rauch:2016pai,Melia:2010bm,Melia:2011gk,Campanario:2013gea,Baglio:2014uba}.
These computations all rely on approximations, while recently the complete NLO corrections have been performed~\cite{Biedermann:2017bss}.
It is therefore interesting to infer in details the quality of the various approximations.
Indeed, apart from Ref.~\cite{Biedermann:2017bss} where it is commented on, \MP{more references?} no detailed comparison of the VBS approximations have been carried out.
Preliminary results have already been made public in %Ref.~\cite{Anders:2018gfr}.

@@

At NLO, the process possesses four contributions of orders $\mathcal{O}{\left(\alpha^{7}\right)}$, $\mathcal{O}{\left(\alphas\alpha^{6}\right)}$, $\mathcal{O}{\left(\alphas^{2}\alpha^{5}\right)}$, and $\mathcal{O}{\left(\alphas^{3}\alpha^{4}\right)}$.
The largest one are the EW corrections~\cite{Biedermann:2017bss,Biedermann:2016yds} of order $\mathcal{O}{\left(\alpha^{7}\right)}$.
The contribution to the order $\mathcal{O}{\left(\alphas\alpha^{6}\right)}$ is the second largest NLO contribution and is often referred to as the QCD corrections to the VBS process.
In the following, this order is the one where our comparisons are focused on.
In this article we will refer to it as simply \emph{NLO}.
As for the LO study, the various predictions are compared at the level of the cross section and differential distributions.
In particular, this allows to infer the accuracy of the so-called VBS approximation, which we will define in more detail later.
To our knowledge, such a detailed study was still missing.

>>>>

In the last few years, several leading order (LO) and next-to-leading order (NLO) computations became available for both the VBS process~\cite{Jager:2009xx,Jager:2011ms,Denner:2012dz,Rauch:2016pai} and its QCD-induced irreducible background process~\cite{Rauch:2016pai,Melia:2010bm,Melia:2011gk,Campanario:2013gea,Baglio:2014uba}.
These computations all rely on approximations, while recently the complete NLO corrections have been performed~\cite{Biedermann:2017bss}.
It is therefore interesting to infer in details the quality of the various approximations.
Indeed, apart from Ref.~\cite{Biedermann:2017bss} where it is commented on, \MP{more references?} no detailed comparison of the VBS approximations have been carried out.
Preliminary results have already been made public in Ref.~\cite{Anders:2018gfr}.

@@

At NLO, the process possesses four contributions of orders $\mathcal{O}{\left(\alpha^{7}\right)}$, $\mathcal{O}{\left(\alphas\alpha^{6}\right)}$, $\mathcal{O}{\left(\alphas^{2}\alpha^{5}\right)}$, and $\mathcal{O}{\left(\alphas^{3}\alpha^{4}\right)}$.
The largest ones are the EW corrections~\cite{Biedermann:2017bss,Biedermann:2016yds} of order $\mathcal{O}{\left(\alpha^{7}\right)}$.
The contribution to the order $\mathcal{O}{\left(\alphas\alpha^{6}\right)}$ is the second largest NLO contribution and is often referred to as the QCD corrections to the VBS process.
In the following, this order is the one where our comparisons are focused on.
In this article we will refer to it as simply \emph{NLO}.
As for the LO study, the various predictions are compared at the level of the cross section and differential distributions.
In particular, this allows to infer the accuracy of the so-called VBS approximation, which we will define in more detail later.
To our knowledge, such a detailed study was still missing.

### 2018-02-07 Draft/scanLO.tex
Minor changes on draft, mostly typos

As explained previously, the low di-jet invariant mass and low jet rapidity separation is dominated by tri-bosons production.
Therefore a comparison of the various approximation in an \emph{inclusive} phase-space volume.
To that end, we have chosen for the inclusive fiducial volume to take the following values for the VBS cuts

\begin{equation}
\label{eq:inclusive}
	m_{jj} > 200\GeV\,,\qquad |\Delta y_{jj}| > 2 .
\end{equation}

The figure is Fig.~\ref{fig:ratio2d_LO}, the ratio of double-differential cross sections in the plan $\left(m_{\Pj\Pj}, \Delta y_{\Pj\Pj}\right)$ is shown.
Two plots are displayed: the ratio of the $|t|^2 + |u|^2$ and $|s|^2 + |t|^2 + |u|^2$ approximations over the full calculation, respectively.
In the first case, the approximation is good within $\pm10\%$ over the whole range apart in the low invariant-mass region at both low and large rapidity difference.
The low rapidity difference region possesses remnants of the tri-bosons contributions.
It is therefore expected that the $|t|^2 + |u|^2$ approximations fails in this region.
The second plot, where the $|s|^2 + |t|^2 + |u|^2$ approximation is considered, displays a better behaviour in the previously mentioned region.
The full calculation is approximated at the level of $\pm10\%$ everywhere apart in the upper left corner where the inclusion of $s$-channel contributions did not improve the approximation.
This is due to ...

\begin{figure}[hbt]
\centering
\includegraphics[scale=0.395]{figures/scanfigures/ratio_tu.pdf}
\includegraphics[scale=0.395]{figures/scanfigures/ratio_stu.pdf}
\caption{Cross sections (fb) per bin in the plan $\left(m_{\Pj\Pj}, \Delta y_{\Pj\Pj}\right)$ at order $\mathcal{O}(\alpha^6)$. 
Ratio of approximated squared amplitudes over the full matrix element. The approximated squared amplitudes are computed as $|\mathcal{A}|^2 \sim |t|^2 + |u|^2$ (left) and $|\mathcal{A}|^2 \sim |s|^2 + |t|^2 + |u|^2$ (right).}
\label{fig:ratio2d_LO}
\end{figure}

>>>>

As explained previously, the low di-jet invariant mass and low jet rapidity separation is dominated by tri-boson production.
Therefore a comparison of the various approximations in an \emph{inclusive} phase-space volume is performed.
To that end, we have chosen for the inclusive fiducial volume to take the following values for the VBS cuts
\begin{equation}
\label{eq:inclusive}
	m_{jj} > 200\GeV\,,\qquad |\Delta y_{jj}| > 2 .
\end{equation}

The figure showing this is Fig.~\ref{fig:ratio2d_LO}, the ratio of double-differential cross sections in the plan $\left(m_{\Pj\Pj}, \Delta y_{\Pj\Pj}\right)$.
Two plots are displayed: the ratio of the $|t|^2 + |u|^2$ and $|s|^2 + |t|^2 + |u|^2$ approximations over the full calculation, respectively.
In the first case, the approximation is good within $\pm10\%$ over the whole range apart in the low invariant-mass region at both low and large rapidity difference.
The low rapidity difference region possesses remnants of the tri-bosons contributions.
It is therefore expected that the $|t|^2 + |u|^2$ approximation fails in this region.
The second plot, where the $|s|^2 + |t|^2 + |u|^2$ approximation is considered, displays a better behaviour in the previously mentioned region.
The full calculation is approximated at the level of $\pm10\%$ everywhere apart in the upper left corner where the inclusion of $s$-channel contributions did not improve the approximation.
This is due to ...

\begin{figure}[hbt]
\centering
\includegraphics[scale=0.395]{figures/scanfigures/ratio_tu.pdf}
\includegraphics[scale=0.395]{figures/scanfigures/ratio_stu.pdf}
\caption{Cross sections (fb) per bin in the plan $\left(m_{\Pj\Pj}, \Delta y_{\Pj\Pj}\right)$ at order $\mathcal{O}(\alpha^6)$.
Ratio of approximated squared amplitudes over the full matrix element. The approximated squared amplitudes are computed as $|\mathcal{A}|^2 \sim |t|^2 + |u|^2$ (top) and $|\mathcal{A}|^2 \sim |s|^2 + |t|^2 + |u|^2$ (bottom).}
\label{fig:ratio2d_LO}
\end{figure}

### 2018-02-07 Draft/scanNLO.tex
Minor changes on draft, mostly typos

Concerning leptonic kinematics, we show in Fig.~\ref{fig:mjjdyjj_1d_3} the distributions of the lepton-lepton invariant mass and of the Zeppenfeld variable of the electron \MP{We should defined the Zep var. somewhere}.
\begin{figure}[hbt]
\centering
{\includegraphics[scale=0.35]{figures/scanfigures/mll_nlo.pdf}}
{\includegraphics[scale=0.35]{figures/scanfigures/zel_nlo.pdf}}
\caption{NLO QCD inclusive study: distributions of the lepton--lepton invariant mass and the electron Zeppenfeld variable, obtained with full ({\sc Recola}) and approximated ({\sc VBFNLO, Bonsay}) amplitudes at order $\mathcal{O}(\alphas\alpha^6)$.} \label{fig:mjjdyjj_1d_3}
\end{figure}
The {\sc VBFNLO} result for the $e^+\mu^+$ invariant mass agrees pretty well with the full curve, generated with {\sc Recola}; the {\sc Bonsay} curve is about 10 \% smaller. The discrepancies are roughly constant over the whole specrum.\\
Instead, the right panel of fig.~\ref{fig:mjjdyjj_1d_3} clearly shows that the Zeppenfeld variable of the positron $z_e$ is strongly affected by the exclusion of $s$--channels, with increasing discrepancy w.r.t. the full result at large values. The muon variable $z_{\mu}$ behaviour is almost identical to $z_e$.\\

>>>>

Concerning leptonic kinematics, we show in Fig.~\ref{fig:mjjdyjj_1d_3} the distributions of the lepton-lepton invariant mass and of the Zeppenfeld variable of the electron \MP{We should defined the Zep var. somewhere}.
\begin{figure}[hbt]
\centering
{\includegraphics[scale=0.35]{figures/scanfigures/mll_nlo.pdf}}
{\includegraphics[scale=0.35]{figures/scanfigures/zel_nlo.pdf}}
\caption{NLO QCD inclusive study: distributions of the lepton--lepton invariant mass and the electron Zeppenfeld variable, obtained with full ({\sc Recola}) and approximated ({\sc VBFNLO, Bonsay}) amplitudes at order $\mathcal{O}(\alphas\alpha^6)$.} \label{fig:mjjdyjj_1d_3}
\end{figure}
The {\sc VBFNLO} result for the $e^+\mu^+$ invariant mass agrees pretty well with the full curve, generated with {\sc Recola}; the {\sc Bonsay} curve is about 10 \% smaller. The discrepancies are roughly constant over the whole specrum.\\
Instead, the right panel of Fig.~\ref{fig:mjjdyjj_1d_3} clearly shows that the Zeppenfeld variable of the positron $z_e$ is strongly affected by the exclusion of $s$--channels, with increasing discrepancy w.r.t. the full result at large values. The muon variable $z_{\mu}$ behaviour is almost identical to $z_e$.\\

### 2018-02-07 Draft/shower.tex
Minor changes on draft, mostly typos

A slightly different setup has been employed for {\sc MG5\_aMC} in order to simplify the calculation: instead of generating the full
${\rm p}{\rm p}\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}$ process, since it is anyway dominated by doubly-resonant contribution, the
events are produced for the process with two stable W$^+$ bosons (${\rm p}{\rm p}\to{\rm W^+}{\rm W^+}{\rm j}{\rm j}$), and these W$^+$ bosons
are decayed with {\sc MadSpin}~\cite{Artoisenet:2012st} (keeping spin correlations) before the PS. Since {\sc MadSpin}\ computes
the partial and total decay width of the W bosons at LO accuracy only, while in Section~\ref{subsec:inputpar} the NLO width is employed,
a small effect (6\%) on the normalisation of distribution is induced. Finally, when the renormalisation
and factorisation scales are set, the $\Delta R_{\Pj\Pl}$ cut is not imposed during the jet-clustering procedure, but this has no visible effect
on the results.
%
\begin{table}[h!]
    \centering
    \begin{tabular}{c|c|c|c}
        Code  &  $\sigma[\rm{fb}]$  \\
        \hline
        {\sc MG5\_aMC}+{\sc Pythia8}&  $1.450 (1.368) \pm 0.$  \\
        {\sc MG5\_aMC}+{\sc Herwig++}&  $1.445 (1.363) \pm 0.$  \\
        {\sc POWHEG}  &  $1.3633 \pm 0.0004$  \\
        {\sc VBFNLO}  &  $1.339 \pm 0.$  \\
        \hline
        {\sc MG5\_aMC}+{\sc Pythia8} (LO)&  $1.352 (1.275) \pm 0.$  \\
        {\sc MG5\_aMC}+{\sc Herwig++} (LO)&  $1.343 (1.267)  \pm 0.$  \\
        {\sc PHANTOM}+{\sc Pythia8} &  $1.235  \pm 0.$  \\
        {\sc PHANTOM}+{\sc Herwig++} &  $1.260  \pm 0.$  \\
    \end{tabular}
    \caption{\label{tab:PSrates} Rates at NLO-QCD (LO-QCD) accuracy matched to parton shower within VBS cuts obtained with the different codes used in this comparison,
    for the ${\rm p}{\rm p}\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}$ process. Numbers in parentheses for the {\sc MG5\_aMC} simulations
    are rescaled to account for the effect related to the boson widths computed by {\sc MadSpin}, see the text for details.}
{\bf MZ: MC uncertainties???}
\end{table}
%
We now turn to present the results of predictions matched to parton shower: the total rates within VBS cuts are displayed in in Table~\ref{tab:PSrates}, both at LO and NLO 
accuracy. For {\sc MG5\_aMC}, the number in parentheses are rescaled to take into account the width effects described in the above paragraph. Once this effect is taken into
account, total rates from different tools agree within few percents. Larger discrepancies however will appear for differential observables, which we are going to discuss in 
the following. For any observable, results will be presented in two plots, shown side-by-side. In the plot on the left (right), (N)LO+PS predictions are shown 
with different colours in the main frame. In the inset, these predictions are compared in both cases with the fixed-order prediciton at NLO accuracy (obtained with 
{\sc VBFNLO}). For the differential observables, the {\sc MG5\_aMC} predictions are \emph{not} rescaled to compensate for the width effect mentioned above.\\

@@

{\sc MG5\_aMC}, while the smallest by {\sc Powheg}, both matched to {\sc Pythia8}. Despite the fact that the same parton shower is employed, the way emissions are treated

@@

Finally, focusing on the third jet, we conclude the list of differential observable with the rapidity and the $z$ variable, Figures~\ref{fig:PSy3} and~\ref{fig:PSz3}. 

>>>>

A slightly different setup has been employed for {\sc MG5\_aMC} in order to simplify the calculation: instead of generating the full
${\rm p}{\rm p}\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}$ process, since it is anyway dominated by doubly-resonant contribution, the
events are produced for the process with two stable W$^+$ bosons (${\rm p}{\rm p}\to{\rm W^+}{\rm W^+}{\rm j}{\rm j}$), and these W$^+$ bosons
are decayed with {\sc MadSpin}~\cite{Artoisenet:2012st} (keeping spin correlations) before the PS. Since {\sc MadSpin}\ computes
the partial and total decay width of the W bosons at LO accuracy only, while in Section~\ref{subsec:inputpar} the NLO width is employed,
a small effect (6\%) on the normalisation of distribution is induced. Finally, when the renormalisation
and factorisation scales are set, the $\Delta R_{\Pj\Pl}$ cut is not imposed during the jet-clustering procedure, but this has no visible effect
on the results.

\begin{table}[h!]
    \centering
    \begin{tabular}{c|c|c|c}
        Code  &  $\sigma[\rm{fb}]$  \\
        \hline
        {\sc MG5\_aMC}+{\sc Pythia8}&  $1.450 (1.368) \pm 0.$  \\
        {\sc MG5\_aMC}+{\sc Herwig++}&  $1.445 (1.363) \pm 0.$  \\
        {\sc POWHEG}  &  $1.3633 \pm 0.0004$  \\
        {\sc VBFNLO}  &  $1.339 \pm 0.$  \\
        \hline
        {\sc MG5\_aMC}+{\sc Pythia8} (LO)&  $1.352 (1.275) \pm 0.$  \\
        {\sc MG5\_aMC}+{\sc Herwig++} (LO)&  $1.343 (1.267)  \pm 0.$  \\
        {\sc PHANTOM}+{\sc Pythia8} &  $1.235  \pm 0.$  \\
        {\sc PHANTOM}+{\sc Herwig++} &  $1.260  \pm 0.$  \\
    \end{tabular}
    \caption{\label{tab:PSrates} Rates at NLO-QCD (LO-QCD) accuracy matched to parton shower within VBS cuts obtained with the different codes used in this comparison,
    for the ${\rm p}{\rm p}\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}$ process. Numbers in parentheses for the {\sc MG5\_aMC} simulations
    are rescaled to account for the effect related to the boson widths computed by {\sc MadSpin}, see the text for details.}
{\bf MZ: MC uncertainties???}
\end{table}

We now turn to present the results of predictions matched to parton shower: the total rates within VBS cuts are displayed in Table~\ref{tab:PSrates}, both at LO and NLO
accuracy. For {\sc MG5\_aMC}, the numbers in parentheses are rescaled to take into account the width effects described in the above paragraph. Once this effect is taken into
account, total rates from different tools agree within few percents. Larger discrepancies however will appear for differential observables, which we are going to discuss in
the following. For any observable, results will be presented in two plots, shown side-by-side. In the plot on the left (right), (N)LO+PS predictions are shown
with different colours in the main frame. In the inset, these predictions are compared in both cases with the fixed-order prediciton at NLO accuracy (obtained with
{\sc VBFNLO}). For the differential observables, the {\sc MG5\_aMC} predictions are \emph{not} rescaled to compensate for the width effect mentioned above.\\

@@

{\sc MG5\_aMC}, while the smallest rate is predicted by {\sc Powheg}, both matched to {\sc Pythia8}. Despite the fact that the same parton shower is employed, the way emissions are treated

@@

Finally, focusing on the third jet, we conclude the list of differential observables with the rapidity and the $z$ variable, Figures~\ref{fig:PSy3} and~\ref{fig:PSz3}.
